---
title: "CCA tissue enhancer analysis: Preliminary analysis"
output: 
  html_notebook:
    toc: true
    toc_float: true

---








# read data

Read samplesheet, read peaksets.  binding matrix is based on peak data
```{r include=F}

rm(list=ls())
library(DiffBind, warn.conflicts=F, quietly=T)
source('C:/Users/cherny/Google Drive/CCA-epigenetics/CCA_enhancers_project/analysis_cca_enhancers_functions.R')
K27ac_enhancers_dba <- dba(sampleSheet="diffbind_samplesheet_K27ac_enhancers.csv")
K27ac_enhancers_dba 
```




Peak overlap stats: Enhancers
```{r}

# enhancers 
dba.overlap(K27ac_enhancers_dba, mode=DBA_OLAP_RATE) 
dba.overlap(K27ac_enhancers_dba, K27ac_enhancers_dba$masks$Tumor, mode=DBA_OLAP_RATE) 
dba.overlap(K27ac_enhancers_dba, K27ac_enhancers_dba$masks$Normal, mode=DBA_OLAP_RATE)

```



Consensus peaks widths
```{r fig.width=12, fig.asp=.3}
summary(K27ac_enhancers_dba$binding[,3] - K27ac_enhancers_dba$binding[,2])

### Enhancer peaks
# individual samples
boxplot(x=lapply(K27ac_enhancers_dba$peaks, function(x){x[,3] - x[,2]}), log="y", main="Enhancer peaks, individual samples", ylab="Peak width", names=K27ac_enhancers_dba$samples$SampleID, cex.axis=.8, las=2)
# merged peaks
boxplot(x=K27ac_enhancers_dba$binding[,3] - K27ac_enhancers_dba$binding[,2], log="y", main="Enhancer peaks, merged", ylab="Peak width", cex.axis=.8, las=2)

```






Recount. Notes:  
Cannot recount followed by minoverlap filter! Because after recount, all samples will have the same peakset.  
Should not increase min_overlap to minimize batch effect, because I will miss out on sites specific to groups with fewer samples (eg. the Normals). 





```{r include=F}
### Recount enhancers
rm(list=ls())
library(DiffBind, warn.conflicts=F, quietly=T)
K27ac_enhancers_dba <- dba(sampleSheet="diffbind_samplesheet_K27ac_enhancers.csv")

K27ac_enhancers_dba_all = dba.count(K27ac_enhancers_dba, fragmentSize=150, minOverlap=2, summits=200) # x sites
save(K27ac_enhancers_dba, K27ac_enhancers_dba_all,
    file="C:/Users/cherny/Google Drive/CCA-epigenetics/CCA_enhancers_project/Rdata_cca_K27ac_0_prelim_recount_enhancers.Rdata")
```


Check consensus peaks width after recount
```{r fig.width=12, fig.asp=.3}
summary(K27ac_promoters_dba_all$binding[,3] - K27ac_promoters_dba_all$binding[,2])
summary(K27ac_enhancers_dba_all$binding[,3] - K27ac_enhancers_dba_all$binding[,2])
summary(K27ac_super_dba_all$binding[,3] - K27ac_super_dba_all$binding[,2])


### Enhancer peaks
# individual samples
boxplot(x=lapply(K27ac_enhancers_dba_all$peaks, function(x){x[,3] - x[,2]}), log="y", main="Enhancer peaks, individual samples", ylab="Peak width", names=K27ac_enhancers_dba_all$samples$SampleID, cex.axis=.8, las=2)
# merged peaks
boxplot(x=K27ac_enhancers_dba_all$binding[,3] - K27ac_enhancers_dba_all$binding[,2], log="y", main="Enhancer peaks, merged", ylab="Peak width", cex.axis=.8, las=2)


```







# plot heatmap & PCA. exclude liver samples for clustering/pca
```{r fig.width=10, fig.asp=.5}

K27ac_enhancers_dba_excLiver = dba(DBA=K27ac_enhancers_dba_all, mask=!K27ac_enhancers_dba_all$masks$Normal_liver, minOverlap = 2) # 71 samples,   sites

K27ac_enhancers_dba_excLiver
dba.plotHeatmap(K27ac_enhancers_dba_excLiver, distMethod="pearson")
dba.plotHeatmap(K27ac_enhancers_dba_excLiver, distMethod="spearman")


# get clinical features
aa_idx = which(as.character(dba.show(K27ac_enhancers_dba_excLiver)$ID) %in% samples_data[grep("Yes",samples_data$AA), "Sample ID"])
fluke_idx = which(as.character(dba.show(K27ac_enhancers_dba_excLiver)$ID) %in% samples_data[samples_data$Fluke=="Fluke-Pos", "Sample ID"])
viral_idx = which(as.character(dba.show(K27ac_enhancers_dba_excLiver)$ID) %in% samples_data[samples_data$HBV=="Positive" | samples_data$HCV=="Positive", "Sample ID"])
intra_idx = which(as.character(dba.show(K27ac_enhancers_dba_excLiver)$ID) %in% samples_data[samples_data$`Anatomical subtype`=="Intrahepatic", "Sample ID"])
peri_idx = which(as.character(dba.show(K27ac_enhancers_dba_excLiver)$ID) %in% samples_data[samples_data$`Anatomical subtype`=="Perihilar", "Sample ID"])
distal_idx = which(as.character(dba.show(K27ac_enhancers_dba_excLiver)$ID) %in% samples_data[samples_data$`Anatomical subtype`=="Distal", "Sample ID"])
extra_idx = which(as.character(dba.show(K27ac_enhancers_dba_excLiver)$ID) %in% samples_data[samples_data$`Anatomical subtype`=="Extrahepatic", "Sample ID"])
cystic_idx = which(as.character(dba.show(K27ac_enhancers_dba_excLiver)$ID) %in% samples_data[samples_data$`Anatomical subtype`=="Cystic duct", "Sample ID"])
col_clus = rep("grey", length(as.character(dba.show(K27ac_enhancers_dba_excLiver)$ID)))
col_clus[as.character(dba.show(K27ac_enhancers_dba_excLiver)$Tissue)=="Normal"] = "black"
col_clus[as.character(dba.show(K27ac_enhancers_dba_excLiver)$Tissue)=="Normal_liver"] = "brown"
col_clus[as.character(dba.show(K27ac_enhancers_dba_excLiver)$ID) %in% samples_data[samples_data$iCluster_infer=="1", "Sample ID"]] = "blue"
col_clus[as.character(dba.show(K27ac_enhancers_dba_excLiver)$ID) %in% samples_data[samples_data$iCluster_infer=="2", "Sample ID"]] = "orange"
col_clus[as.character(dba.show(K27ac_enhancers_dba_excLiver)$ID) %in% samples_data[samples_data$iCluster_infer=="3", "Sample ID"]] = "green"
col_clus[as.character(dba.show(K27ac_enhancers_dba_excLiver)$ID) %in% samples_data[samples_data$iCluster_infer=="2/3", "Sample ID"]] = "yellow"
col_clus[as.character(dba.show(K27ac_enhancers_dba_excLiver)$ID) %in% samples_data[samples_data$iCluster_infer=="4", "Sample ID"]] = "red"



# pca: rank transform
K27ac_enhancers_excLiver_signals_rank = apply(K27ac_enhancers_excLiver_signals, 2, function(x){rank(x, ties.method = "average")})
K27ac_enhancers_excLiver_signals_rank_pca = prcomp(t(K27ac_enhancers_excLiver_signals_rank), center=T, scale=T)
# plot pca
layout(mat=matrix(c(1,2), 1, 2))
plot(x=K27ac_enhancers_excLiver_signals_rank_pca$x[,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[,2],
     col=col_clus, pch=16)
plot(x=K27ac_enhancers_excLiver_signals_rank_pca$x[,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[,3],
     col=col_clus, pch=16)
# plot pca with risk factors
layout(mat=matrix(c(1,2), 1, 2))
plot(x=K27ac_enhancers_excLiver_signals_rank_pca$x[,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[,2],
     col=col_clus, pch=16)
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[aa_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[aa_idx,2], pch=2, cex=2.5, col="gray30")
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[fluke_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[fluke_idx,2], pch=1, cex=2.5, col="gray30")
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[viral_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[viral_idx,2], pch="V", cex=1, col="gray30")
plot(x=K27ac_enhancers_excLiver_signals_rank_pca$x[,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[,3],
     col=col_clus, pch=16)
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[aa_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[aa_idx,3], pch=2, cex=2.5, col="gray30")
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[fluke_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[fluke_idx,3], pch=1, cex=2.5, col="gray30")
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[viral_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[viral_idx,3], pch="V", cex=1, col="gray30")
# plot pca with subtype
layout(mat=matrix(c(1,2), 1, 2))
plot(x=K27ac_enhancers_excLiver_signals_rank_pca$x[,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[,2],
     col=col_clus, pch=16)
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[peri_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[peri_idx,2], pch="p", cex=.9, col="gray30")
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[distal_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[distal_idx,2], pch="d", cex=.9, col="gray30")
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[extra_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[extra_idx,2], pch="e", cex=.9, col="gray30")
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[cystic_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[cystic_idx,2], pch="c", cex=.9, col="gray30")
plot(x=K27ac_enhancers_excLiver_signals_rank_pca$x[,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[,3],
     col=col_clus, pch=16)
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[peri_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[peri_idx,3], pch="p", cex=.9, col="gray30")
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[distal_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[distal_idx,3], pch="d", cex=.9, col="gray30")
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[extra_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[extra_idx,3], pch="e", cex=.9, col="gray30")
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[cystic_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[cystic_idx,3], pch="c", cex=.9, col="gray30")
# plot pca with genotype
layout(mat=matrix(c(1,2), 1, 2))
plot(x=K27ac_enhancers_excLiver_signals_rank_pca$x[,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[,2],
     col=col_clus, pch=16)
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[bap1_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[bap1_idx,2], pch="b", cex=.9, col="gray30")
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[bap1trunc_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[bap1trunc_idx,2], pch="B", cex=.9, col="gray30")
plot(x=K27ac_enhancers_excLiver_signals_rank_pca$x[,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[,3],
     col=col_clus, pch=16)
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[bap1_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[bap1_idx,3], pch="b", cex=.9, col="gray30")
points(x=K27ac_enhancers_excLiver_signals_rank_pca$x[bap1trunc_idx,1], y=K27ac_enhancers_excLiver_signals_rank_pca$x[bap1trunc_idx,3], pch="B", cex=.9, col="gray30")



```




















```{r include=F}
options(scipen=999)



########### Save the dba objects for downstream analysis ###
# save the original peaksets for downstream analysis
K27ac_dba_origPeaks = dba(DBA=K27ac_enhancers_dba)
K27ac_origPeaks = K27ac_dba_origPeaks$peaks
names(K27ac_origPeaks) = as.character(dba.show(K27ac_dba_origPeaks)$ID)
K27ac_origPeaks = lapply(K27ac_origPeaks, function(x){
  data.frame(chr=x$V1, start=x$V2, end=x$V3, stringsAsFactors = F)
})

# save enhancers and enh+sup DBAs
K27ac_dba = K27ac_enhancers_dba_all 
K27ac_dba$contrasts = NULL

# output deseq2 normalization factors for signal tracks normalization
deseq2facs_all_enh = get_deseq2_normfacs(K27ac_enhancers_dba_all)
write.table(deseq2facs_all_enh, file="20190521_cca_K27ac_Deseq2NormFacs.txt", quote=F, row.names=T, sep="\t", col.names=F)


save(K27ac_dba, K27ac_origPeaks, file="C:/Users/cherny/Google Drive/CCA-epigenetics/CCA_enhancers_project/Rdata_cca_k27ac_0_prelim_final.Rdata")






```




